Fail Faster

### 2025-03-15 paper/paper.tex
continue to write splitmix section

%  The precise
% details of how SplitMix works are unimportant for the present paper, but the
% main operations are defined in terms of bitwise operations on 64-bit integers.
% However,

% This has a significant impact on the performance of generators. A single call to
% one of the \bq library functions may sample from \texttt{Splittable\_random}
% multiple times, and each sample from \texttt{Splittable\_random} allocates 9
% words \jwc{this is a call to \texttt{next\_int64}}.  While seemingly small, we
% will see in Section~\ref{section:eval} that this amount of allocation has a
% significant impact on significant.

@@

\subsection{Fast Random Number Generators}
\label{subsection:faster-rng}

\jwc{
  \begin{itemize}
    \item Using the fastest RNG possible is important.
    \item In OCaml, the splittalbe random RNG is not as fast as it could be, because of the Ints issue.
    \item In this section, we describe how to resolve the ints issue by calling out ot the FFI. 
    \item All of this can be done with Jane Street bleeding-edge compiler extensions, though these are not yet compatible with MetaOCaml.
    \item Because different libraries 
  \end{itemize}
}

\section{Evaluation}
\subsection{Implementing and Testing Generators}
\label{subsection:pbt-for-pbt}
\jwc{Talk about the difftesting to ensure staging maintains semantics, and also different RNGs have identical semantics.}

\subsection{Benchmarking speed \& resource usage}

\jwc{NOTE: we should test generator speed across both languages, but speed -> bugfinding ability in only OCaml.}
\jwc{
  Baseline generators to test speed in both languages:
  \begin{itemize}
    \item Single int
    \item Pair of ints, constrained
    \item List of ints
  \end{itemize}
}
\subsection{Impact on bug-finding ability}
\jwc{Staging type-derived generators (which are the most commonly used ones) can turn a timeout into a found bug!}
\section{Conclusion \& Future Work}
\section{Related Work}
\jwc{A generator is a parser of randomness, and people have implemented lots of staged parser libraries!}
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{bib}

%%
%% If your work has an appendix, this is the place to put it.
\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.

>>>>

@@

\subsection{Faster SplitMix with Unboxed Int}
\label{subsection:faster-rng}

\jwc{Unclear what we should call this section}

As we discussed in Section~\ref{section:motiv}, inefficient RNG choice is
another bottleneck for finding bugs fast. While generator libraries by
and large use sensible RNG algorithms, they are not explicitly chosen
with performance in mind. Indeed, much faster RNGs \cn and fast sources
of entropy \cn exist. \jwc{... more here?}
To demonstrate that a faster RNG can significanty impact bugfinding power, we
use the natural experiment provided by OCaml's inefficient implementation of
SplitMix. By replacing this slow RNG with a faster but extensionally equivalent
implemetnation, we are able to precisely quantify the bugfinding speedup that using a RNG gives across a range of PBT scenarios.

We emphasize that while we believe that the insight that faster RNG translates to faster bug finding
is a cross-langauge one, the specific technical contents of this section are OCaml-specific.
In most other PBT frameworks \cn, the RNG operates on by machine
integers, so this \emph{particular} inefficiency does not exist.

The precise details of how the SplitMix algorithm works are unimportant for the
present paper, but the critical component is that all of its operations are are
defined in terms of arithmetic bitwise operations on 64-bit integers.  In OCaml,
a
% however,

% this has a significant impact on the performance of generators. a single call to
% one of the \bq library functions may sample from \texttt{splittable\_random}
% multiple times, and each sample from \texttt{splittable\_random} allocates 9
% words \jwc{this is a call to \texttt{next\_int64}}.  while seemingly small, we
% will see in section~\ref{section:eval} that this amount of allocation has a
% significant impact on significant.

\jwc{
  \begin{itemize}
    \item as we discussed in section~\ref{section:motiv}, 
    \item faster rngs and sources of entropy exist, though they are extensionally inequivalent. 
    \item 
    \item 
    \item here's how splitmix is slow: the core of splitmix is nextint64, which does allocations for each operation.
    \item if we could use actual int64s, we wouldn't have to allocate.
    \item so we do this:
    \begin{itemize}
      \item we implement splitmix in c (which uses normal int64s) and real int64 arithmetic, and use the ocaml ffi to call out to the c functions. t
      \item this could be done with jane street bleeding-edge compiler extensions, though these are not yet implemented in a a version of the ocaml compiler that also has staging.
    \end{itemize}
  \end{itemize}
}

\section{evaluation}
\subsection{implementing and testing generators}
\label{subsection:pbt-for-pbt}
\jwc{talk about the difftesting to ensure staging maintains semantics, and also different rngs have identical semantics.}

\subsection{benchmarking speed \& resource usage}

\jwc{note: we should test generator speed across both languages, but speed -> bugfinding ability in only ocaml.}
\jwc{
  baseline generators to test speed in both languages:
  \begin{itemize}
    \item single int
    \item pair of ints, constrained
    \item list of ints without bind (use map): lots of sampling, minimal binds.
    \item unableled trees of a fixed size (no weighted union) minimal sampling, lots of binds.
  \end{itemize}
}
\subsection{impact on bug-finding ability}
\jwc{staging type-derived generators (which are the most commonly used ones) can turn a timeout into a found bug!}
\section{conclusion \& future work}
\section{related work}
\jwc{a generator is a parser of randomness, and people have implemented lots of staged parser libraries!}
%%
%% the next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{acm-reference-format}
\bibliography{bib}

%%
%% if your work has an appendix, this is the place to put it.
\end{document}
\endinput
%%
%% end of file `sample-sigplan.tex'.
